---
title: "Module 8 exercise: Fitting Basic Statistical Models"
author: "Tzu-Chun Chu"
date: "10/12/2021"
output: html_document
---

# load required packages
```{r setup, warning=F}
# Load needed packages
library(tidymodels) 
library(tidyverse)
library(magrittr)
library(rlang)
library(dotwhisker)  # for visualizing regression results
```

# Load data
```{r data}
processdata <- readRDS(here::here("data","processed_data","processeddata.rds"))

```


# Model fitting
```{r}
### Linear Linear Model ###

#First, let's set engine for performing linear regression
linear_fit <- 
	linear_reg() %>% 
	set_engine("lm")
```


# Linear Model 1: regression body temperature on runny nose 
```{r}
lm_1 <- 
	linear_fit %>% 
	fit(BodyTemp ~ RunnyNose, data = processdata)
lm_1
```


```{r}
#summarize the results
tidy(lm_1)
```


# Linear Model 2: regression body temperature on all predictors
```{r}
lm_2 <- linear_fit %>% 
	fit(BodyTemp ~ ., data = processdata)
lm_2
```


```{r}
#summarize the results
tidy(lm_2)

```

# plotting
```{r}
#generate a dot-and-whisker plot of linear regression results with all predictors
dotwhisker_cont_all_pred <- tidy(lm_2) %>% 
	dotwhisker::dwplot(dot_args = list(size = 2, color = "black"),
										 whisker_args = list(color = "black"),
										 vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```


```{r}
#Compares the model results for the linear regression model with just the main predictor and all predictors.
##Analysis of deviance and model selection
res_lm_aov <-
	anova(lm_1$fit, lm_2$fit, test = "Chisq")
res_lm_aov
```

The p-value is smaller than 0.05, therefore the model with all predictors had better fit than the single variable model

# Logistic regression Model 
```{r}
# Let's set engine for performing linear regression
logistic_fit <- 
	logistic_reg() %>% 
	set_mode("classification") %>% 
	set_engine("glm")
```


# Logistic Model 1: regression body temperature on runny nose 
```{r}
glm_1 <- 
	logistic_fit %>% 
	fit(Nausea ~ RunnyNose, data = processdata)
glm_1
```

```{r}
#summarize the results
tidy(glm_1)
```

# Logistic Model 2: regression body temperature on all predictors
```{r}
glm_2 <- logistic_fit %>% 
	fit(Nausea ~ ., data = processdata)
glm_2
```


```{r}
#summarize the results
tidy(glm_2)
```

# Plotting
```{r}
#generate a dot-and-whisker plot of logistic regression results with all predictors
dotwhisker_binary_all_pred <- tidy(glm_2) %>% 
	dotwhisker::dwplot(dot_args = list(size = 2, color = "black"),
										 whisker_args = list(color = "black"),
										 vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))

```

```{r}
#Compares the model results for the logistic regression model with just the main predictor and all predictors.
##Analysis of deviance and model selection
res_glm_aov <-
	anova(glm_1$fit, glm_2$fit, test = "Chisq")
res_glm_aov
```

The p-value is smaller than 0.05, therefore the model with all predictors had better fit than the single variable model